Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-27 article.tex
Introduce the difference in preprocessing

This section is a complete description of the method specifically tuned for the
challenge. Basically, it is an optimized version of the method presented in
Sections~\ref{sec:filter}~to~\ref{sec:results}, therefore only differences
between both version of our method will be highlighted.

@@

As we introduce a composed function in section~\ref{sec:filter}, let us
describe a new composed function with optimized filters to improve the quality
of the inference process.

Besides filters $f_1$ and $f_2$ respectively defined by equations \ref{eq
:symetric-median} and \ref{eq:weighted-asymetric-median}, we also consider two
other low-pass filters:

\begin{align}
f_3(x^t_i) &= x^{t-1}_i + x^{t}_i + x^{t+1}_i + x^{t+2}_i \label{eq:asymetric-median-forward}, \\
f_4(x^t_i) &=  x_i^t + x^{t+1}_i  + x^{t+2}_i + x^{t+3}_i  .
\label{eq:asymetric-median}
\end{align}

Theses filters lead, taken individually, to worse performances  than filters
$f_1$ and $f_2$ but the averaging approach over all low-pass filters gets
better for the ROC AUC.

\begin{table}[ht]\label{tab:comparison}
%\caption{Evaluation of different methods on the four \textit{normal} datasets\label{tab:comparison}}
\centering
\small
\begin{tabular}{| l | c c c c | c c c c |}
\hline
& \multicolumn{4}{c|}{AUROC} & \multicolumn{4}{c|}{AUPRC} \\
\textit{Averaging} $\backslash$ \textit{normal-} & \textit{1} & \textit{2} & \textit{3} & \textit{4} & \textit{1} & \textit{2} & \textit{3} & \textit{4} \\
\hline
\hline
 with $f_1$, $f_2$ & 0.937 & 0.935 & 0.935 & 0.931 & 0.391 &  \textbf{0.390} &  0.385 & \textbf{0.375}  \\
 with $f_1$, $f_2$, $f_3$, $f_4$ & \textbf{0.938} & \textbf{0.936} & \textbf{0.936} & \textbf{0.932} & 0.391 & 0.389 & 0.385 & 0.374\\
\hline
\end{tabular}
\caption{Performance on \textit{normal-1,2,3,4} with partial correlation with different averaging approaches.}
\end{table}

As a computation trick, it appears that using a parametrized hard-threshold
filter slightly improves performances. Introducing the parameter $c$ in
function $h$ yields to an optimized filter
\begin{align}
h^*(x^t_i) = (x_i^t)^c \text{ } \mathbb{1}(x_i^t \ge \tau) \text{ with } \tau > 0,
\end{align}
where $c$ is set to $0.9$ after experiments.

@@

We report here experiments on particular datasets provided by the organizers
which are similar to \textit{normal} datasets, i.e., made of 1000 neurons,
except for one parameter. We compute, as done in Section~\ref{sec:results},
both AUROC and AUPRC and results are reported in
Table~\ref{tab:results_appendix}. We will briefly recall what characterizes
each datasets:
\begin{description}
\item[lowcon]: Similar network but with a lower number of connections per neuron on average.
\item[highcon]: Similar network but with a higher number of connections per neuron on average.
\item[lowcc]: Similar network but with a lower clustering coefficient on average.
\item[highcc]: Similar network but with a higher clustering coefficient on average.
\item[normal-3-highrate]: Same topology as \textit{normal-3} but with an higher firing frequency, i.e., with highly active neurons.
\item[normal-4-lownoise]: Same topology as \textit{normal-4} but with a better signal to noise ratio.
\end{description}

>>>>

This section provides a full description of the method specifically tuned for
the challenge. Since it is no more than an optimized version of the method
presented in Sections~\ref{sec:filter}~to~\ref{sec:results}, the description
will be limited to explaining the differences that exist between this method
and its optimized version.

@@

In Section~\ref{sec:filter}, we introduce two composed filtering functions in
order (i) to remove noise induced by light scattering artifact, (ii) to take
into account the slow decay of fluorescence signal and (iii) to reduce the
importance of global high activity in the network. However, in order to win the
challenge, we have heavily optimized several different, but yet similar,
filtering composed function. Taken alone, those composed functions in
conjunction with partial correlation might lead to worse AUROC, but improved
whenever they are averaged.

Besides filters $f_1$ and $f_2$ respectively defined by equations \ref{eq
:symetric-median} and \ref{eq:weighted-asymetric-median}, we also consider two
other low-pass filters
\begin{align}
f_3(x^t_i) &= x^{t-1}_i + x^{t}_i + x^{t+1}_i + x^{t+2}_i, \label{eq:asymetric-median-forward} \\
f_4(x^t_i) &=  x_i^t + x^{t+1}_i  + x^{t+2}_i + x^{t+3}_i. \label{eq:asymetric-median}
\end{align}

Theses filters lead, taken individually, to worse performances  than filters
$f_1$ and $f_2$ but averaging the results of all low-pass filters leads to better ROC AUC.
\begin{table}[ht]\label{tab:comparison}
%\caption{Evaluation of different methods on the four \textit{normal} datasets\label{tab:comparison}}
\centering
\small
\begin{tabular}{| l | c c c c | c c c c |}
\hline
& \multicolumn{4}{c|}{AUROC} & \multicolumn{4}{c|}{AUPRC} \\
\textit{Averaging} $\backslash$ \textit{normal-} & \textit{1} & \textit{2} & \textit{3} & \textit{4} & \textit{1} & \textit{2} & \textit{3} & \textit{4} \\
\hline
\hline
 with $f_1$, $f_2$ & 0.937 & 0.935 & 0.935 & 0.931 & 0.391 &  \textbf{0.390} &  0.385 & \textbf{0.375}  \\
 with $f_1$, $f_2$, $f_3$, $f_4$ & \textbf{0.938} & \textbf{0.936} & \textbf{0.936} & \textbf{0.932} & 0.391 & 0.389 & 0.385 & 0.374\\
\hline
\end{tabular}
\caption{Performance on \textit{normal-1,2,3,4} with partial correlation with different averaging approaches.}
\end{table}

As computation trick, it appears that using a parametrized hard-threshold
filter slightly improves performances. Introducing the parameter $c$ in
function $h$ yields an optimized filter
\begin{align}
h^*(x^t_i) = (x_i^t)^c \text{ } \mathbb{1}(x_i^t \ge \tau) \text{ with } \tau > 0,
\end{align}
where $c$ is set to $0.9$ after experiments.

@@

We report here experiments on particular datasets provided by the organizers
which are similar to \textit{normal} datasets, i.e., made of 1000 neurons,
except for one parameter. We compute, as done in Section~\ref{sec:results},
both AUROC and AUPRC and results are reported in
Table~\ref{tab:results_appendix}. We will briefly recall what characterizes
each dataset:
\begin{description}
\item[lowcon]: Similar network but with a lower number of connections per neuron on average.
\item[highcon]: Similar network but with a higher number of connections per neuron on average.
\item[lowcc]: Similar network but with a lower clustering coefficient on average.
\item[highcc]: Similar network but with a higher clustering coefficient on average.
\item[normal-3-highrate]: Same topology as \textit{normal-3} but with a higher firing frequency, i.e., with highly active neurons.
\item[normal-4-lownoise]: Same topology as \textit{normal-4} but with a better signal to noise ratio.
\end{description}
